---
title: "Stack Overflow"
description: |
  In this article, we deal with the unbalanced dataset in classification. 
author:
  - name: Thinh Tran
    url: {}
date: 09-12-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE, dpi = 180, 
                      fig.width = 8, fig.height = 5)
```

# Prerequisite
```{r}
library(tidyverse)
theme_set(theme_light())
```

# Prepare the data 
```{r}
stack_overflow <- read_csv("https://raw.githubusercontent.com/juliasilge/supervised-ML-case-studies-course/master/data/stack_overflow.csv") %>% 
  mutate(remote = factor(remote, levels = c("Remote", "Not remote"))) %>% # reorder level for consistency
  mutate(across(where(is.character), factor)) %>%  # all char type to factor type for modelling purpose
  mutate(across(where(is.logical), as.numeric))
# Overview
head(stack_overflow)
```

## EDA
###  Dimension
```{r}
stack_overflow %>% 
  glimpse()
```


### Distribution of some interested variables
```{r}
stack_overflow %>% 
  count(remote, sort = TRUE)

stack_overflow %>% 
  count(country, sort= TRUE)

stack_overflow %>% 
  count(years_coded_job, sort = TRUE)

```

### Making assumption with the relationship between variables
```{r}
pivot_longer_stack <- stack_overflow %>% 
  pivot_longer(data_scientist:web_developer, 
               names_to = "job") %>%
  filter(value == 1) %>% 
  mutate(job = str_to_title(str_replace_all(job, "_", " "))) 
  
pivot_longer_stack %>% 
  count(job, remote, sort = TRUE) %>% 
  mutate(job = fct_reorder(job, n)) %>% 
  ggplot(aes(job, n, fill = remote)) + 
  geom_col() + 
  coord_flip() + 
  labs(
    x = "Job Title",
    y = "# Respondents",
    fill = "Is Remote"
  ) 
```


```{r}
stack_overflow %>% 
  ggplot(aes(remote, years_coded_job, fill = remote)) + 
  geom_boxplot(alpha = .5) +
  labs(
    x = "",
    y = "# Years",
    fill = "Is Remote",
    title = "Distribution of Professional Coding Experience Between Remote Status"
  )
```


# Predict whether a respondent is working remotely or not with classification
```{r}
stack_overflow %>% 
  count(remote, sort = TRUE)

```

There are about ten times more non-remote developers than there are remote developers. We are dealing with unbalanced datasets. 

```{r}
# load tidymodels 
library(tidymodels)
library(themis)

# create stack_select dataset 
stack_select <- stack_overflow %>% 
  select(-respondent) # redundant in the model

# split data into training and testing tests
set.seed(1234)
stack_split <- stack_select %>% 
  initial_split(p = 0.8, stata = remote)

stack_train <- training(stack_split)
stack_test <- testing(stack_split)

```


```{r}
# define the roles of all columns in the dataset
stack_recipe <- recipe(remote ~ ., data = stack_train) %>% 
  step_downsample(remote) 

# pre-processing
stack_prep <- prep(stack_recipe)
stack_down <- juice(stack_prep)

# check on the result of downsample
stack_down %>% 
  count(remote)
```

## Predicting remote status

### Methods: Logistic Regression and Decision Tree

Prepare the specification for fitting the data

```{r}
glm_spec <- parsnip::logistic_reg() %>% 
  set_engine("glm")

tree_spec <- parsnip::decision_tree() %>% 
  set_engine("rpart") %>% 
  set_mode("classification")
```

Build a workflow to keep tracks with the process of training model
```{r}
# build a workflow

stack_wf <- workflow() %>% 
  add_recipe(stack_recipe) 

stack_glm <- stack_wf %>% 
  add_model(glm_spec) %>% 
  fit(data = stack_train)

stack_tree <- stack_wf %>% 
  add_model(tree_spec) %>% 
  fit(data = stack_train)

stack_glm 
stack_tree
```

### Compare the results
```{r}
results <- stack_test %>%
    bind_cols(predict(stack_glm, stack_test) %>%
                  rename(.pred_glm = .pred_class)) %>%
    bind_cols(predict(stack_tree, stack_test) %>%
                  rename(.pred_tree = .pred_class))

results %>%
    conf_mat(truth = remote, estimate = .pred_glm)

results %>%
    conf_mat(truth = remote, estimate = .pred_tree)

## Calculate accuracy
accuracy(results, truth = remote, estimate = .pred_glm)
accuracy(results, truth = remote, estimate = .pred_tree)

## Calculate positive predict value
ppv(results, truth = remote, estimate = .pred_glm)
ppv(results, truth = remote, estimate = .pred_tree)
```


