---
title: "Voter"
description: |
  Introduction to Cross Validation
author:
  - name: Thinh Tran
    url: {}
date: 09-12-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prerequisite
```{r}
library(tidyverse)
library(gt)
theme_set(theme_light())
```


# Loading the data
```{r}
voters <- read_csv("https://raw.githubusercontent.com/juliasilge/supervised-ML-case-studies-course/master/data/voters.csv")
```

Useful info

```{r}
tribble(~Response, ~ Code,
        "strongly agree", 1,
        "agree", 2,
        "disagree", 3,
        "strongly disagree", 4) %>% 
  mutate(Response = str_to_title(Response)) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_column_labels(everything())
  ) %>% 
  opt_table_lines("default")
```

# EDA 

```{r}
voters %>% 
  glimpse()
```

## How many people voted? 
```{r}
voters %>% 
  count(turnout16_2016)
```

## How do the responses on the survey vary with voting behavior?

```{r}
voters %>%
    group_by(turnout16_2016) %>%
    summarise(`Elections don't matter` = mean(RIGGED_SYSTEM_1_2016 <= 2),
              `Economy is getting better` = mean(econtrend_2016 == 1),
              `Crime is very important` = mean(imiss_a_2016 == 2)) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_column_labels(everything())
  ) %>% 
  tab_options(
    column_labels.border.top.color = "white",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3)
  ) %>% 
  tab_spanner(
    label = md("Survey questions"),
    columns = 2:4) %>% 
  cols_label(
    turnout16_2016 = "Turnout",
    `Elections don't matter` = "Elections don't matter",
    `Economy is getting better` = "Economy is getting better",
    `Crime is very important` = "Crime is very important"
  ) %>% 
  fmt_percent(
    columns = 2:4
  )
```


```{r}
voters %>%
    ggplot(aes(econtrend_2016, after_stat(density), fill = turnout16_2016)) +
    geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) +
    labs(title = "How is the economy among the voter view?")
```

# Make a prediction on whether a person vote or don't vote based on their answers

## Training and testing data
```{r}
voters_select <- voters %>% 
  mutate(turnout16_2016 = factor(turnout16_2016)) %>% 
  select(-case_identifier)

# loading tidymodels 
library(tidymodels)
library(themis)

# split the data into training and testing sets
voters_split <- voters_select %>% 
  initial_split(prop = .8, stata = turnout16_2016)

voters_train <- training(voters_split)
voters_test <- testing(voters_split)

```

We implement upsampling, we add more of the people who did not vote (just more of the same ones we already have) until the proportion is equal and the classes are balanced.
## Create cross-validation folds

```{r}
vote_folds <- rsample::vfold_cv(voters_train, v = 10, repeats = 5)

```

## Preprocess with a recipe 
```{r}
voters_recipe <- recipe(turnout16_2016 ~ ., data = voters_train) %>%
  step_upsample(turnout16_2016)

```

## Creating a modelling workflow

```{r}
# specify a logistic model 
glm_spec <- logistic_reg() %>% 
  set_engine("glm")

vote_wf <- workflow() %>% 
  add_recipe(voters_recipe) %>% 
  add_model(glm_spec)

set.seed(234)
glm_res <- vote_wf %>%
  fit_resamples(
    vote_folds, 
    metrics = metric_set(roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
  )
glimpse(glm_res)
  
```

```{r}
## specify a ranger model 
rf_spec <- rand_forest() %>% 
  set_engine("ranger") %>% 
  set_mode("classification")

# Add the recipe + model to a workflow 

vote_wf <- workflow() %>% 
  add_recipe(voters_recipe) %>% 
  add_model(rf_spec)

set.seed(234)
rf_res <- vote_wf %>%
    fit_resamples(
        vote_folds,
        metrics = metric_set(roc_auc, sens, spec),
        control = control_resamples(save_pred = TRUE))

glimpse(rf_res)
```

## Performance metrics from resampling

```{r}
collect_metrics(glm_res)
collect_metrics(rf_res)
```


## Testing    

```{r}
## Final fit
vote_final <- vote_wf %>%
    last_fit(voters_split)

## Confusion matrix
vote_final %>% 
    collect_predictions() %>% 
    conf_mat(turnout16_2016, .pred_class)
```

